High-dimensional Asymptotics of Information Transfer for Linear Regression under Covariate Shifts

### 2022-02-13 prior_versions/revision_journal/conclusions.tex
update

\section{Discussion and conclusions}\label{sec_conclude}

This work studied generalization properties of a widely used hard parameter sharing approach for multi-task learning.
We provided sharp bias-variance tradeoffs of HPS in high-dimensional linear regression.
Using these results, we analyzed how varying sample sizes and covariate shifts impact HPS, and rigorously explained several empirical phenomena such as negative transfer and covariate shift related to these dataset properties.
We validated our theory and conducted further studies on text classification tasks.
We describe several open questions for future work.
%First, it would be interesting to tighten our estimate in Corollary \ref{cor_MTL_loss}, which would extend the observation in Figure \ref{fig_size} to small $n_1$.
%Second, it would be interesting to extend our result to classification problems such as logistic regression.
First, our result in Corollary \ref{cor_MTL_loss} involves an error term that scales down with $n_1$.
Tightening this error bound requires showing the limit of $\normFro{({Z^{(1)}}^{\top} Z^{(1)} + {Z^{(2)}}^{\top} Z^{(2)})^{-1} {Z^{(1)}}^{\top} Z^{(1)}}^2$ for two isotropic sample covariance matrices.
This requires studying the asymptotic singular values distribution of the non-symmetric matrix $({Z^{(1)}}^{\top} Z^{(1)})^{-1}{Z^{(2)}}^{\top} Z^{(2)}+\id$, which is still an open problem in random matrix theory.
%FY: $+\id$ is very important and makes the problem very hard; otherwise the problem can be solved with current RMT methods.
The eigenvalue distribution of this matrix, which has been obtained in \cite{Fmatrix}, might be helpful towards resolving this problem.
%but its singular values will follow a different distribution since the matrix is not symmetric.
 %might require new techniques beyond the current ones in random matrix theory .%\HZ{to add}.
Second, it would be interesting to extend our results to classification problems.
Several recent work have made remarkable progress for logistic regression in the high-dimensional setting, e.g. \cite{sur2019modern}.
An interesting question is to study logistic regression in a multiple-sample setting.

>>>>

\section{Conclusions}\label{sec_conclude}

Distribution shift is a fundamental challenge in applying transfer learning.
This work formulated a theoretical setup in which questions related to the transfer effect can be formally analyzed.
The setup can reproduce several interesting phenomena in the context of transfer learning.
Precise asympotics for the risk of a two-layer linear neural network are shown under various kinds of distribution shift.
Such analysis has led to a number of theoretical insights and practical implications for applying transfer learning to text classification tasks.

%We describe several open questions for future work.
%First, it would be interesting to tighten our estimate in Corollary \ref{cor_MTL_loss}, which would extend the observation in Figure \ref{fig_size} to small $n_1$.
%Second, it would be interesting to extend our result to classification problems such as logistic regression.
It is a very interesting question to derive the asymptotic limit under general covariate and model shift.
%This requires showing the limit of $\normFro{\big(({Z^{(1)}})^{\top} Z^{(1)} + {Z^{(2)}}^{\top} Z^{(2)}\big)^{-1} {Z^{(1)}}^{\top} Z^{(1)}}^2$ for two sample covariance matrices.
We remark that likely such a result will require studying the asymptotic distribution of the singular values of asymmetric matrices, which is technically challenging.
%FY: $+\id$ is very important and makes the problem very hard; otherwise the problem can be solved with current RMT methods.
%The eigenvalue distribution of this matrix, which has been obtained in \cite{Fmatrix}, might be helpful towards resolving this problem.
%but its singular values will follow a different distribution since the matrix is not symmetric.
 %might require new techniques beyond the current ones in random matrix theory .%\HZ{to add}.
Another interesting question is to study the impact of distribution shift in other settings such as logistic regression \cite{sur2019modern}.

### 2022-02-13 prior_versions/revision_journal/manuscript_bk.tex
update

\title{High-dimensional Asymptotics of Transferring from Different Domains in Linear Regression}

\runtitle{High-dimensional Asymptotics of Transferring from Different Domains}
%\thankstext{T1}{A sample additional note to the title.}

@@

\section{Main Results and Proof Techniques}\label{sec_HPS}
In this section, we define our model for the setting of transfer learning, and introduce several transfer learning estimators that will be considered in this paper.

@@

\subsection{The case of isotropic covariances} \label{sec_sizeratio}

@@

\textbf{Variance asymptotic limit.}
In Theorem \ref{main_cor}, we will show that for $z$ in a small neighborhood around $0$, when $p$ goes to infinity, $G(z)$ converges to the following limit
\be \label{defn_piw}
\Gi(z) \define \begin{pmatrix} (a_{1}(z)\Lambda^2  +  (a_{2}(z)- z)\id_{p\times p})^{-1} & 0 & 0 \\ 0 & - \frac{n_1+n_2}{n_1} a_{1}(z)\id_{n_1\times n_1} & 0 \\ 0 & 0 & -\frac{n_1+n_2}{n_2}a_{2}(z)\id_{n_2\times n_2}  \end{pmatrix},\ee
where $a_1(z)$ and $a_2(z)$ are the unique solutions to the following self-consistent equations
\be\label{selfomega_a}
\begin{split}
&a_1(z) + a_2(z) = 1 - \frac{1}{n_1 + n_2} \bigbrace{\sum_{i=1}^p \frac{\lambda_i^2 a_1(z) + a_2(z)}{\lambda_i^2 a_1(z) + a_2(z) - z}}, \\ %
&a_1(z) + \frac{1}{n_1 + n_2}\bigbrace{\sum_{i=1}^p \frac{\lambda_i^2 a_1(z)}{\lambda_i^2 a_1(z) + a_2(z) - z}} = \frac{n_1}{n_1 + n_2}.
\end{split}
\ee
The existence and uniqueness of solutions to the above system are shown in Lemma \ref{lem_mbehaviorw}.
Given this result, we now show that when $z = 0$, the matrix limit $\Gi(0)$ implies the variance limit shown in equation \eqref{lem_cov_shift_eq}.
First, we have that $a_1 = a_1(0)$ and $a_2 = a_2(0)$ since the equations in \eqref{selfomega_a} reduce to equation \eqref{eq_a12extra} when $z=0$.
Second, since $W^{-1}$ is the upper-left block matrix of $G(0)$, we have that $W^{-1}$ converges to $ (a_1\Lambda^2 + a_2\id)^{-1} $.
Using the fact that $\tr[\Sigma^{(2)} \hat{\Sigma}^{-1}] = (n_1 + n_2)^{-1}\bigtr{W^{-1}} $, we get that when $p$ goes to infinity, %
\begin{align*}
\bigtr{\Sigma^{(2)} \hat{\Sigma}} \rightarrow \frac{1}{n_1+n_2}\bigtr{(a_1 \Lambda^2 + a_2\id)^{-1}} &= \frac1{n_1+n_2}\bigtr{(a_1 M^{\top}M + a_2 \id)^{-1}} \\
&=\frac{1}{n_1+n_2} \bigtr{\Sigma^{(2)} (a_1 \Sigma^{(1)} + a_2 \Sigma^{(2)})^{-1}},
\end{align*}
where we note that $M^\top M = (\Sigma^{(2)})^{-1/2} \Sigma^{(1)} (\Sigma^{(2)})^{-1/2}$ and its SVD is equal to $V^{\top}\Lambda^2 V$.

\textbf{Bias asymptotic limit.}
For the bias limit in equation \eqref{lem_cov_derv_eq}, we show that it is governed by the derivative of $(W - z\id)^2$ with respect to $z$ at $z = 0$.
First, we can express the empirical bias term in equation \eqref{lem_cov_derv_eq} as %
\begin{align}\label{calculate G'}
(n_1 + n_2)^2 \hat{\Sigma}^{-1}\Sigma^{(2)}\hat{\Sigma}^{-1} = {\Sigma^{(2)}}^{-1/2} V W^{-2} V^{\top} {\Sigma^{(2)}}^{-1/2}.
\end{align}
Let $\cal G(z):=(W-z\id )^{-1}$ denote the resolvent of $W$.
Our key observation is that $\frac{\dd{\cal G(z)}}{\dd z} =  \cal G^2(z)$.
Hence, provided that the limit of $(W - z\id)^{-1}$ is $(a_1(z) \Lambda^2 + (a_2(z) - z) \id)^{-1}$ near $z = 0$, the limit of $\frac{\dd{\cal G(0)}}{\dd z}$ satisfies
\begin{align}\label{cal G'0}
\frac{\dd \cal G(0)}{\dd z} \to \frac{-\frac{\dd a_1(0)}{\dd z}\Lambda^2 - (\frac{\dd a_2(0)}{\dd z} - 1)\id}{(a_{1}(0)\Lambda^2 + a_{2}(0)\id_p)^2}.
\end{align}
To find the derivatives of $a_1(z)$ and $a_2(z)$, we take the derivatives on both sides of the system of equations \eqref{selfomega_a}.
Let $a_3 = - \frac{\dd a_1(0)}{\dd z}$ and $a_4 = - \frac{\dd a_2(0)}{\dd z}$.
One can verify that $a_3$ and $a_4$ satisfy the self-consistent equations in \eqref{eq_a34extra} (details omitted).
Applying equation \eqref{cal G'0} to equation \eqref{calculate G'}, we obtain the bias limit.

@@

\textbf{Schur complement and self-consistent equations.}
First, we consider the special case where $Z^{(1)}$ and $Z^{(2)}$ are both multivariate Gaussian random matrices.
By rotational invariance, we have that $Z^{(1)} U$ and $Z^{(2)} V$ are still multivariate Gaussian random matrices.
Next, we use the Schur complement formula to deal with the resolvent $G(z)$.
We show that $G(z)$'s diagonal entries satisfy a set of self-consistent equations in the limit, leading to equations in \eqref{selfomega_a}.
On the other hand, $G(z)$'s off-diagonal entries are approximately zero using standard concentration bounds.
Finally, we extend our result to general random matrices under the finite $\varphi$-th moment condition.
We prove an anisotropic local law using recent developments in random matrix theory \cite{erdos2017dynamical,Anisotropic}.
The proof of Theorem \ref{thm_main_RMT} is shown in Appendix \ref{appendix RMT}.

@@

\section{Conclusions}\label{sec_conclude}

This work studied generalization properties of a widely used hard parameter sharing approach for multi-task learning.
We provided sharp bias-variance tradeoffs of HPS in high-dimensional linear regression.
Using these results, we analyzed how varying sample sizes and covariate shifts impact HPS, and rigorously explained several empirical phenomena such as negative transfer and covariate shift related to these dataset properties.
We validated our theory and conducted further studies on text classification tasks.
We describe several open questions for future work.
%First, it would be interesting to tighten our estimate in Corollary \ref{cor_MTL_loss}, which would extend the observation in Figure \ref{fig_size} to small $n_1$.
%Second, it would be interesting to extend our result to classification problems such as logistic regression.
First, our result in Corollary \ref{cor_MTL_loss} involves an error term that scales down with $n_1$.
Tightening this error bound requires showing the limit of $\normFro{({Z^{(1)}}^{\top} Z^{(1)} + {Z^{(2)}}^{\top} Z^{(2)})^{-1} {Z^{(1)}}^{\top} Z^{(1)}}^2$ for two isotropic sample covariance matrices.
This requires studying the asymptotic singular values distribution of the non-symmetric matrix $({Z^{(1)}}^{\top} Z^{(1)})^{-1}{Z^{(2)}}^{\top} Z^{(2)}+\id$, which is still an open problem in random matrix theory.
%FY: $+\id$ is very important and makes the problem very hard; otherwise the problem can be solved with current RMT methods.
The eigenvalue distribution of this matrix, which has been obtained in \cite{Fmatrix}, might be helpful towards resolving this problem.
%but its singular values will follow a different distribution since the matrix is not symmetric.
 %might require new techniques beyond the current ones in random matrix theory .%\HZ{to add}.
Second, it would be interesting to extend our results to classification problems.
Several recent work have made remarkable progress for logistic regression in the high-dimensional setting, e.g. \cite{sur2019modern}.
An interesting question is to study logistic regression in a multiple-sample setting.

\begin{supplement}
\textbf{Supplement to ``High-dimensional Asymptotics of Transferring from Different Domains in Linear Regression"}.
In the Internet Appendix \cite{MTL_suppl}, we provide the proofs of the technical results in Sections \ref{sec_HPS}-\ref{sec_diff}, including Lemma \ref{lem_HPS_loss}, Theorem \ref{thm_many_tasks}, Theorem \ref{cor_MTL_loss}, Proposition \ref{lem_hat_v}, Theorem \ref{thm_main_RMT} and Proposition \ref{prop_main_RMT}.
%and give the technical proofs of the main theorems, Theorems \ref{thm_regularbm}, \ref{thm_twgram}, \ref{thm_twgraph} and  \ref{thm_twgraph_sparse}.
\end{supplement}

>>>>

\title{High-dimensional Asymptotics of Information Transfer for Linear Regression under Covariate Shifts}

\runtitle{High-dimensional Asymptotics of Information Transfer}
%\thankstext{T1}{A sample additional note to the title.}

@@

\section{High-dimensional asymtotics of information transfer}\label{sec_HPS}
In this section, we define our model for the setting of transfer learning, and introduce several transfer learning estimators that will be considered in this paper.

@@

\subsection{A tight result for isotropic covariances} \label{sec_sizeratio}

@@

\paragraph{Variance asymptotic limit.}
In Theorem \ref{main_cor}, we will show that for $z$ in a small neighborhood around $0$, when $p$ goes to infinity, $G(z)$ converges to the following limit
\be \label{defn_piw}
\Gi(z) \define \begin{pmatrix} (a_{1}(z)\Lambda^2  +  (a_{2}(z)- z)\id_{p\times p})^{-1} & 0 & 0 \\ 0 & - \frac{n_1+n_2}{n_1} a_{1}(z)\id_{n_1\times n_1} & 0 \\ 0 & 0 & -\frac{n_1+n_2}{n_2}a_{2}(z)\id_{n_2\times n_2}  \end{pmatrix},\ee
where $a_1(z)$ and $a_2(z)$ are the unique solutions to the following self-consistent equations
\be\label{selfomega_a}
\begin{split}
&a_1(z) + a_2(z) = 1 - \frac{1}{n_1 + n_2} \bigbrace{\sum_{i=1}^p \frac{\lambda_i^2 a_1(z) + a_2(z)}{\lambda_i^2 a_1(z) + a_2(z) - z}}, \\ %
&a_1(z) + \frac{1}{n_1 + n_2}\bigbrace{\sum_{i=1}^p \frac{\lambda_i^2 a_1(z)}{\lambda_i^2 a_1(z) + a_2(z) - z}} = \frac{n_1}{n_1 + n_2}.
\end{split}
\ee
The existence and uniqueness of solutions to the above system are shown in Lemma \ref{lem_mbehaviorw}.
Given this result, we now show that when $z = 0$, the matrix limit $\Gi(0)$ implies the variance limit shown in equation \eqref{lem_cov_shift_eq}.
First, we have that $a_1 = a_1(0)$ and $a_2 = a_2(0)$ since the equations in \eqref{selfomega_a} reduce to equation \eqref{eq_a12extra} when $z=0$.
Second, since $W^{-1}$ is the upper-left block matrix of $G(0)$, we have that $W^{-1}$ converges to $ (a_1\Lambda^2 + a_2\id)^{-1} $.
Using the fact that $\tr[\Sigma^{(2)} \hat{\Sigma}^{-1}] = (n_1 + n_2)^{-1}\bigtr{W^{-1}} $, we get that when $p$ goes to infinity, %
\begin{align*}
\bigtr{\Sigma^{(2)} \hat{\Sigma}} \rightarrow \frac{1}{n_1+n_2}\bigtr{(a_1 \Lambda^2 + a_2\id)^{-1}} &= \frac1{n_1+n_2}\bigtr{(a_1 M^{\top}M + a_2 \id)^{-1}} \\
&=\frac{1}{n_1+n_2} \bigtr{\Sigma^{(2)} (a_1 \Sigma^{(1)} + a_2 \Sigma^{(2)})^{-1}},
\end{align*}
where we note that $M^\top M = (\Sigma^{(2)})^{-1/2} \Sigma^{(1)} (\Sigma^{(2)})^{-1/2}$ and its SVD is equal to $V^{\top}\Lambda^2 V$.

\paragraph{Bias asymptotic limit.}
For the bias limit in equation \eqref{lem_cov_derv_eq}, we show that it is governed by the derivative of $(W - z\id)^2$ with respect to $z$ at $z = 0$.
First, we can express the empirical bias term in equation \eqref{lem_cov_derv_eq} as %
\begin{align}\label{calculate G'}
(n_1 + n_2)^2 \hat{\Sigma}^{-1}\Sigma^{(2)}\hat{\Sigma}^{-1} = {\Sigma^{(2)}}^{-1/2} V W^{-2} V^{\top} {\Sigma^{(2)}}^{-1/2}.
\end{align}
Let $\cal G(z):=(W-z\id )^{-1}$ denote the resolvent of $W$.
Our key observation is that $\frac{\dd{\cal G(z)}}{\dd z} =  \cal G^2(z)$.
Hence, provided that the limit of $(W - z\id)^{-1}$ is $(a_1(z) \Lambda^2 + (a_2(z) - z) \id)^{-1}$ near $z = 0$, the limit of $\frac{\dd{\cal G(0)}}{\dd z}$ satisfies
\begin{align}\label{cal G'0}
\frac{\dd \cal G(0)}{\dd z} \to \frac{-\frac{\dd a_1(0)}{\dd z}\Lambda^2 - (\frac{\dd a_2(0)}{\dd z} - 1)\id}{(a_{1}(0)\Lambda^2 + a_{2}(0)\id_p)^2}.
\end{align}
To find the derivatives of $a_1(z)$ and $a_2(z)$, we take the derivatives on both sides of the system of equations \eqref{selfomega_a}.
Let $a_3 = - \frac{\dd a_1(0)}{\dd z}$ and $a_4 = - \frac{\dd a_2(0)}{\dd z}$.
One can verify that $a_3$ and $a_4$ satisfy the self-consistent equations in \eqref{eq_a34extra} (details omitted).
Applying equation \eqref{cal G'0} to equation \eqref{calculate G'}, we obtain the bias limit.

@@

\paragraph{Schur complement and self-consistent equations.}
First, we consider the special case where $Z^{(1)}$ and $Z^{(2)}$ are both multivariate Gaussian random matrices.
By rotational invariance, we have that $Z^{(1)} U$ and $Z^{(2)} V$ are still multivariate Gaussian random matrices.
Next, we use the Schur complement formula to deal with the resolvent $G(z)$.
We show that $G(z)$'s diagonal entries satisfy a set of self-consistent equations in the limit, leading to equations in \eqref{selfomega_a}.
On the other hand, $G(z)$'s off-diagonal entries are approximately zero using standard concentration bounds.
Finally, we extend our result to general random matrices under the finite $\varphi$-th moment condition.
We prove an anisotropic local law using recent developments in random matrix theory \cite{erdos2017dynamical,Anisotropic}.
The proof of Theorem \ref{thm_main_RMT} is shown in Appendix \ref{appendix RMT}.

@@

\section{Discussions}\label{sec_conclude}
\FY{Conclusions are not needed for AOS. If you want to have this section, please use it to give some additional remarks that are not discussed yet, and to discuss about possible futuer works.}
This work studied generalization properties of a widely used hard parameter sharing approach for multi-task learning.
We provided sharp bias-variance tradeoffs of HPS in high-dimensional linear regression.
Using these results, we analyzed how varying sample sizes and covariate shifts impact HPS, and rigorously explained several empirical phenomena such as negative transfer and covariate shift related to these dataset properties.
We validated our theory and conducted further studies on text classification tasks.
We describe several open questions for future work.
%First, it would be interesting to tighten our estimate in Corollary \ref{cor_MTL_loss}, which would extend the observation in Figure \ref{fig_size} to small $n_1$.
%Second, it would be interesting to extend our result to classification problems such as logistic regression.
First, our result in Corollary \ref{cor_MTL_loss} involves an error term that scales down with $n_1$.
Tightening this error bound requires showing the limit of $\normFro{({Z^{(1)}}^{\top} Z^{(1)} + {Z^{(2)}}^{\top} Z^{(2)})^{-1} {Z^{(1)}}^{\top} Z^{(1)}}^2$ for two isotropic sample covariance matrices.
This requires studying the asymptotic singular values distribution of the non-symmetric matrix $({Z^{(1)}}^{\top} Z^{(1)})^{-1}{Z^{(2)}}^{\top} Z^{(2)}+\id$, which is still an open problem in random matrix theory.
%FY: $+\id$ is very important and makes the problem very hard; otherwise the problem can be solved with current RMT methods.
The eigenvalue distribution of this matrix, which has been obtained in \cite{Fmatrix}, might be helpful towards resolving this problem.
%but its singular values will follow a different distribution since the matrix is not symmetric.
 %might require new techniques beyond the current ones in random matrix theory .%\HZ{to add}.
Second, it would be interesting to extend our results to classification problems.
Several recent work have made remarkable progress for logistic regression in the high-dimensional setting, e.g. \cite{sur2019modern}.
An interesting question is to study logistic regression in a multiple-sample setting.

\begin{supplement}
\textbf{Supplement to ``High-dimensional Asymptotics of Information Transfer for Linear Regression under Covariate Shifts"}.
In the Internet Appendix \cite{MTL_suppl}, we provide the proofs of the technical results in Sections \ref{sec_HPS}-\ref{sec_diff}, including Lemma \ref{lem_HPS_loss}, Theorem \ref{thm_many_tasks}, Theorem \ref{cor_MTL_loss}, Proposition \ref{lem_hat_v}, Theorem \ref{thm_main_RMT} and Proposition \ref{prop_main_RMT}.
%and give the technical proofs of the main theorems, Theorems \ref{thm_regularbm}, \ref{thm_twgram}, \ref{thm_twgraph} and  \ref{thm_twgraph_sparse}.
\end{supplement}

### 2022-02-13 prior_versions/revision_journal/results.tex
update

We show precise estimates of the bias and variance of HPS and almost sharp convergence rates.
In Section \ref{sec3_cov}, we characterize the variance of HPS under covariate shift, which extends the classical result from Lemma \ref{fact_tr} to the sum of two sample covariance matrices.
We provide a detailed analysis of the impact of covariate shift on the excess risk of HPS.
In Section \ref{sec_sizeratio}, we characterize the bias of HPS under model shift.
Combined with the variance estimates, we describe three regimes of information transfer in the random-effect model, depending on the sample size $n_1$ and the model shift parameter $\mu$.
Finally, Section \ref{sec3_combined} extends our results to settings with both covariate and model shift.
%On the other hand, in case (ii), we will provide an estimate on the bias term, which becomes exact only when $n_1\gg p$.

@@

\subsection{Covariate and model shift}\label{sec3_combined}

@@

\begin{figure}[!t]
	\begin{subfigure}[b]{0.5\textwidth}
		\centering
		\includegraphics[width=0.8\textwidth]{figures/covariate_and_model_shift_a.eps}
		\caption{Fixed $\mu$, different $\lambda$}
		\label{fig_sec3_cov_mo_a}
	\end{subfigure}
	\begin{subfigure}[b]{0.5\textwidth}
		\centering
		\includegraphics[width=0.8\textwidth]{figures/covariate_and_model_shift_b.eps}
		\caption{Fixed $\lambda$, different $\mu$}
		\label{fig_sec3_cov_mo_b}
	\end{subfigure}
	\caption{We further show that our observations in Figures \ref{fig_sec31} and \ref{fig_sec3_model_shift} extend to settings with both covariate and model shift.
	Both simulation use $p = 100, n_2 = 300, \sigma = 1/2$.
	Figure \ref{fig_sec3_cov_mo_a} fixes $\mu = 0.1$ while varying $\lambda, n_1$.
	Figure \ref{fig_sec3_cov_mo_b} fixes $\lambda = 4$ while varying $\mu, n_1$.}
	\label{fig_sec33}
\end{figure}

@@

\begin{theorem}[Bias estimates under covariate and model shift]\label{prop_main_RMT}
%The bias equation \eqref{Lbias} satisfies the following limit with high probability: Let $S$ be an arbitrary subset of the unit sphere in dimension $p$ whose size is polynomial in $p$, for any unit vector $w\in S$,
Under Assumption \ref{assm_big1}, for any small constant $c>0$ and large constant $C>0$, there exists a high probability event $\Xi$, on which the following estimate  holds for $L_{\bias}(a)$ in \eqref{Lbias}:
			\begin{align}
				& \bigabs{ L_{\bias}(a) -   (\beta^{(1)}- a\beta^{(2)} )^\top (\Sigma^{(1)})^{1/2} \Pi(a)(\Sigma^{(1)})^{1/2} (\beta^{(1)}- a\beta^{(2)})   }  \nonumber\\
				& \le \left[\left( 1+\sqrt{\frac{p}{n_1}}\right)^4 - 1 +\OO\left( n_1^{-1/2+2/\varphi + c}\right)\right] \frac{n_1^2 \lambda_1^2 \left\|(\Sigma^{(1)})^{1/2} \left(\beta^{(1)}- a\beta^{(2)}\right) \right\|^2}{  [(\sqrt{n_1}-\sqrt{p})^2 \lambda_p^2+ (\sqrt{n_2}-\sqrt{p})^2]^2}  \nonumber\\
				& + p^{-C} \left[\| \beta^{(1)}\|^2 + \| \beta^{(2)}\|^2 \right], \label{lem_cov_derv_eq}
			\end{align}
%			\begin{align}\label{lem_cov_derv_eq}
%				\bigabs{ L_{\bias}(a) - \left\|\frac{n_1}{n_1+n_2}  \frac{\left[a_3 M(a)^\top M(a) + (a_4 + 1) \right]^{1/2}}{ a_1 M(a)^\top M(a) + a_2 } M(a)^\top (\Sigma^{(1)})^{1/2} \left(\beta^{(1)}- a\beta^{(2)}\right)\right\|^2   } \le  \frac{p^{-c_{\varphi}}}{(n_1+n_2)^2},
%			\end{align}
				uniformly in all $a\in \R$. Here $\lambda_1$ and $\lambda_p$ are respectively the largest and smallest singular values of $M(a)$, $\Pi(a)$ is a $p\times p$ matrix defined as
				$$\Pi(a):=\frac{n_1^2}{(n_1+n_2)^2}  M(a)  \frac{a_3 M(a)^\top M(a) + (a_4 + 1) }{[a_1 M(a)^\top M(a) + a_2 ]^2} M(a)^\top  ,$$
				 and $(a_{3},a_4)$ is the solution of the following system of equations % with $b_k = \frac1{p}\sum_{i=1}^p \frac{\lambda_i^{2k}} {(\lambda_i^2 a_1 + a_2)^2}$, for $k = 0, 1, 2$:
		\be  \label{eq_a34extra}
		\begin{split}
				& a_3 + a_4 = \frac{1}{n_1 + n_2}\sum_{i=1}^p \frac{1}{\lambda_i^2 a_1 + a_2}, \\
				& a_3 + \frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{\lambda_i^2 (a_2 a_3-a_1 a_4 )}{(\lambda_i^2 a_1 + a_2)^2} = \frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{\lambda_i^2 a_1}{(\lambda_i^2 a_1 + a_2)^{2}},
%				\left(\frac{\rho_1}{a_1^{2}} -  b_2  \right)\cdot  a_3 -  b_1 \cdot  a_4 = b_1,\quad \left(\frac{\rho_2}{a_2^{2}}-  b_0\right)\cdot  a_4 - b_1 \cdot  a_3
%				= b_0.
			\end{split}
			\ee
			where we recall that $(a_1,a_2)$ is the solution of \eqref{eq_a12extra}.
\end{theorem}

>>>>

We show precise estimates of the bias and variance of HPS and almost sharp convergence rates.
In Section \ref{sec3_cov}, we characterize the variance of HPS under covariate shift, which extends the classical result from Lemma \ref{fact_tr} to the sum of two sample covariance matrices with covariate shift.
Based on the characterization, We provide a detailed analysis of the impact of covariate shift on the excess risk of HPS.
In Section \ref{sec_sizeratio}, we characterize the bias of HPS under model shift.
Combined with the variance estimates, we describe three regimes of information transfer in the random-effect model, depending on the sample size $n_1$ and the model shift parameter $\mu$.
Finally, Section \ref{sec3_combined} extends these results to settings under covariate and model shifts.
%On the other hand, in case (ii), we will provide an estimate on the bias term, which becomes exact only when $n_1\gg p$.

@@

\subsection{Covariate and model shifts}\label{sec3_combined}
\todo{write down the formal statement}

@@

\begin{figure}[!t]
	\begin{subfigure}[b]{0.5\textwidth}
		\centering
		\includegraphics[width=0.8\textwidth]{figures/covariate_and_model_shift_a.eps}
		\caption{Fixed $\mu$, different $\lambda$}
		\label{fig_sec3_cov_mo_a}
	\end{subfigure}
	\begin{subfigure}[b]{0.5\textwidth}
		\centering
		\includegraphics[width=0.8\textwidth]{figures/covariate_and_model_shift_b.eps}
		\caption{Fixed $\lambda$, different $\mu$}
		\label{fig_sec3_cov_mo_b}
	\end{subfigure}
	\caption{We further show that our observations in Figures \ref{fig_sec31} and \ref{fig_sec3_model_shift} extend to settings under covariate and model shifts.
	Both simulation use $p = 100, n_2 = 300, \sigma = 1/2$.
	Figure \ref{fig_sec3_cov_mo_a} fixes $\mu = 0.1$ while varying $\lambda, n_1$.
	Figure \ref{fig_sec3_cov_mo_b} fixes $\lambda = 4$ while varying $\mu, n_1$.}
	\label{fig_sec33}
\end{figure}

@@

\begin{theorem}[Bias estimates under covariate and model shifts]\label{prop_main_RMT}
%The bias equation \eqref{Lbias} satisfies the following limit with high probability: Let $S$ be an arbitrary subset of the unit sphere in dimension $p$ whose size is polynomial in $p$, for any unit vector $w\in S$,
Under Assumption \ref{assm_big1}, for any small constant $c>0$ and large constant $C>0$, there exists a high probability event $\Xi$, on which the following estimate  holds for $L_{\bias}(a)$ in \eqref{Lbias}:
			\begin{align}
				& \bigabs{ L_{\bias}(a) -   (\beta^{(1)}- a\beta^{(2)} )^\top (\Sigma^{(1)})^{1/2} \Pi(a)(\Sigma^{(1)})^{1/2} (\beta^{(1)}- a\beta^{(2)})   }  \nonumber\\
				& \le \left[\left( 1+\sqrt{\frac{p}{n_1}}\right)^4 - 1 +\OO\left( n_1^{-1/2+2/\varphi + c}\right)\right] \frac{n_1^2 \lambda_1^2 \left\|(\Sigma^{(1)})^{1/2} \left(\beta^{(1)}- a\beta^{(2)}\right) \right\|^2}{  [(\sqrt{n_1}-\sqrt{p})^2 \lambda_p^2+ (\sqrt{n_2}-\sqrt{p})^2]^2}  \nonumber\\
				& + p^{-C} \left[\| \beta^{(1)}\|^2 + \| \beta^{(2)}\|^2 \right], \label{lem_cov_derv_eq}
			\end{align}
%			\begin{align}\label{lem_cov_derv_eq}
%				\bigabs{ L_{\bias}(a) - \left\|\frac{n_1}{n_1+n_2}  \frac{\left[a_3 M(a)^\top M(a) + (a_4 + 1) \right]^{1/2}}{ a_1 M(a)^\top M(a) + a_2 } M(a)^\top (\Sigma^{(1)})^{1/2} \left(\beta^{(1)}- a\beta^{(2)}\right)\right\|^2   } \le  \frac{p^{-c_{\varphi}}}{(n_1+n_2)^2},
%			\end{align}
				uniformly in all $a\in \R$. Here $\lambda_1$ and $\lambda_p$ are respectively the largest and smallest singular values of $M(a)$, $\Pi(a)$ is a $p\times p$ matrix defined as
				$$\Pi(a):=\frac{n_1^2}{(n_1+n_2)^2}  M(a)  \frac{a_3 M(a)^\top M(a) + (a_4 + 1) }{[a_1 M(a)^\top M(a) + a_2 ]^2} M(a)^\top  ,$$
				 and $(a_{3},a_4)$ is the solution of the following system of equations % with $b_k = \frac1{p}\sum_{i=1}^p \frac{\lambda_i^{2k}} {(\lambda_i^2 a_1 + a_2)^2}$, for $k = 0, 1, 2$:
		\be  \label{eq_a34extra}
		\begin{split}
				& a_3 + a_4 = \frac{1}{n_1 + n_2}\sum_{i=1}^p \frac{1}{\lambda_i^2 a_1 + a_2}, \\
				& a_3 + \frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{\lambda_i^2 (a_2 a_3-a_1 a_4 )}{(\lambda_i^2 a_1 + a_2)^2} = \frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{\lambda_i^2 a_1}{(\lambda_i^2 a_1 + a_2)^{2}},
%				\left(\frac{\rho_1}{a_1^{2}} -  b_2  \right)\cdot  a_3 -  b_1 \cdot  a_4 = b_1,\quad \left(\frac{\rho_2}{a_2^{2}}-  b_0\right)\cdot  a_4 - b_1 \cdot  a_3
%				= b_0.
			\end{split}
			\ee
			where we recall that $(a_1,a_2)$ is the solution of \eqref{eq_a12extra}.
\end{theorem}
\todo{consider changing $a_3, a_4$ to $\alpha_3, \alpha_4$ because of conflict with $a$}
